---
title: "GSE75010_mediation"
author: "Kyle Abraham Campbell"
date: "8/14/2021"
output: html_document
---

```{r setup, include=FALSE}
#devtools::install_github("BS1125/CMAverse")
library(CMAverse)
library(here)
library(limma)
library(mediation)
library(tidymodels)
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```

Load files from GSE75010_dex_analysis.Rmd
```{r}
# Outdated datasets
#df <- readRDS(file = here("data", "2022-04-12_med_data.rda"))
#df.all <- readRDS(file = here("data", "2022-04-12_all_med_data.rda"))

df <- readRDS(file = here("data", "2022-05-19_med_data.rda"))
df.all <- readRDS(file = here("data", "2022-05-19_all_med_data.rda"))
```

# g-formula Mediation in CMAVerse {.tabset}

```{r}
# Determining value of m for controlled effect
m.mean.control <- df %>%
  dplyr::filter(phenotype == "Control") %>%
  dplyr::summarise(mean = mean(Dim.3)) %>%
  as.numeric()

m.mean.preeclampsia <- df %>%
  dplyr::filter(phenotype == "Preeclampsia") %>%
  dplyr::summarise(mean = mean(Dim.3)) %>%
  as.numeric()

find_m <- function(var) {
  
  m.mean.control <- df %>%
  dplyr::filter(phenotype == "Control") %>%
  dplyr::summarise(mean = mean(.data[[var]])) %>%
  as.numeric()

  m.mean.preeclampsia <- df %>%
  dplyr::filter(phenotype == "Preeclampsia") %>%
  dplyr::summarise(mean = mean(.data[[var]])) %>%
  as.numeric(m.mean.preeclampsia)

  return(c(m.mean.control, m.mean.preeclampsia))
}

find_m("Dim.1")
find_m("Dim.2")
find_m("Dim.3")
find_m("Dim.4")
find_m("Dim.5")

m.mean.control <- find_m("Dim.3")[1]
m.mean.preeclampsia <- find_m("Dim.3")[2]

```

## PC3 Interaction, m = mean(control)
Using the g-formula approach because it allows for continuous mediator
```{r, results='hide'}
res_gformula.m.astar <- cmest(data = df, model = "gformula", outcome = "FLT1", exposure = "phenotype",
                      mediator = "Dim.3", basec = c("sex", "batch"), postc = "ga", EMint = TRUE,
                      mreg = list("linear"), yreg = "linear", postcreg = list("linear"),
                      astar = 0, a = 1, mval = list(m.mean.control), 
                      estimation = "imputation", inference = "bootstrap", nboot = 1000)
```

```{r}
summary(res_gformula.m.astar)
```

## PC3 Interaction, m = mean(preeclampsia)
```{r, results='hide'}
res_gformula.m.a <- cmest(data = df, model = "gformula", outcome = "FLT1", exposure = "phenotype",
                      mediator = "Dim.3", basec = c("sex", "batch"), postc = "ga", EMint = TRUE,
                      mreg = list("linear"), yreg = "linear", postcreg = list("linear"),
                      astar = 0, a = 1, mval = list(m.mean.preeclampsia), 
                      estimation = "imputation", inference = "bootstrap", nboot = 1000)
```

```{r}
summary(res_gformula.m.a)
```

## PC3, no interaction, m = mean(control)
```{r, results='hide'}
res_gformula.m.a.noint <- cmest(data = df, model = "gformula", outcome = "FLT1", exposure = "phenotype",
                      mediator = "Dim.3", basec = c("sex", "batch"), postc = "ga", EMint = FALSE,
                      mreg = list("linear"), yreg = "linear", postcreg = list("linear"),
                      astar = 0, a = 1, mval = list(m.mean.control), 
                      estimation = "imputation", inference = "bootstrap", nboot = 1000)
```

```{r}
summary(res_gformula.m.a.noint)
```

## PC3, no interaction, m = mean(preeclampsia)
```{r, results='hide'}
res_gformula.m.a.noint <- cmest(data = df, model = "gformula", outcome = "FLT1", exposure = "phenotype",
                      mediator = "Dim.3", basec = c("sex", "batch"), postc = "ga", EMint = FALSE,
                      mreg = list("linear"), yreg = "linear", postcreg = list("linear"),
                      astar = 0, a = 1, mval = list(m.mean.preeclampsia), 
                      estimation = "imputation", inference = "bootstrap", nboot = 1000)
```

```{r}
summary(res_gformula.m.a.noint)
```

## Each PC in one model with interaction

### FLT1
Analytic model with interaction
```{r, results='hide'}
res_gformula.m.a <- cmest(data = df, model = "gformula", outcome = "FLT1", exposure = "phenotype",
                      mediator = c("Dim.1", "Dim.2", "Dim.3", "Dim.4", "Dim.5"), basec = c("sex", "batch"), postc = "ga", EMint = TRUE,
                      mreg = list("linear", "linear", "linear", "linear", "linear"), yreg = "linear", postcreg = list("linear"),
                      astar = 0, a = 1, mval = list(0, 0, 0, 0, 0), estimation = "imputation", inference = "bootstrap", nboot = 1000)
```

```{r}
summary(res_gformula.m.a)
```

### LEP
Analytic model with interaction
```{r, results='hide'}
res_gformula.m.a <- cmest(data = df.all, model = "gformula", outcome = "LEP", exposure = "phenotype",
                      mediator = c("Dim.1", "Dim.2", "Dim.3", "Dim.4", "Dim.5"), basec = c("sex", "batch"), postc = "ga", EMint = TRUE,
                      mreg = list("linear", "linear", "linear", "linear", "linear"), yreg = "linear", postcreg = list("linear"),
                      astar = 0, a = 1, mval = list(0, 0, 0, 0, 0), estimation = "imputation", inference = "bootstrap", nboot = 1000)
```

```{r}
summary(res_gformula.m.a)
```

### ENG
Analytic model with interaction
```{r, results='hide'}
res_gformula.m.a <- cmest(data = df.all, model = "gformula", outcome = "ENG", exposure = "phenotype",
                      mediator = c("Dim.1", "Dim.2", "Dim.3", "Dim.4", "Dim.5"), basec = c("sex", "batch"), postc = "ga", EMint = TRUE,
                      mreg = list("linear", "linear", "linear", "linear", "linear"), yreg = "linear", postcreg = list("linear"),
                      astar = 0, a = 1, mval = list(0, 0, 0, 0, 0), estimation = "imputation", inference = "bootstrap", nboot = 1000)
```

```{r}
summary(res_gformula.m.a)
```

## Each PC in one model without interaction
### FLT1
Final model without interaction
```{r, results='hide'}
res_gformula.m.a <- cmest(data = df, model = "gformula", outcome = "FLT1", exposure = "phenotype",
                      mediator = c("Dim.1", "Dim.2", "Dim.3", "Dim.4", "Dim.5"), basec = c("sex", "batch"), postc = "ga", EMint = FALSE,
                      mreg = list("linear", "linear", "linear", "linear", "linear"), yreg = "linear", postcreg = list("linear"),
                      astar = 0, a = 1, mval = list(0, 0, 0, 0, 0), 
                      estimation = "imputation", inference = "bootstrap", nboot = 1000)
```

```{r}
summary(res_gformula.m.a)
```
### LEP
Final model without interaction
```{r, results='hide'}
res_gformula.m.a <- cmest(data = df.all, model = "gformula", outcome = "LEP", exposure = "phenotype",
                      mediator = c("Dim.1", "Dim.2", "Dim.3", "Dim.4", "Dim.5"), basec = c("sex", "batch"), postc = "ga", EMint = FALSE,
                      mreg = list("linear", "linear", "linear", "linear", "linear"), yreg = "linear", postcreg = list("linear"),
                      astar = 0, a = 1, mval = list(0, 0, 0, 0, 0), 
                      estimation = "imputation", inference = "bootstrap", nboot = 1000)
```

```{r}
summary(res_gformula.m.a)
```

### ENG
Final model without interaction
```{r, results='hide'}
res_gformula.m.a <- cmest(data = df.all, model = "gformula", outcome = "ENG", exposure = "phenotype",
                      mediator = c("Dim.1", "Dim.2", "Dim.3", "Dim.4", "Dim.5"), basec = c("sex", "batch"), postc = "ga", EMint = FALSE,
                      mreg = list("linear", "linear", "linear", "linear", "linear"), yreg = "linear", postcreg = list("linear"),
                      astar = 0, a = 1, mval = list(0, 0, 0, 0, 0), 
                      estimation = "imputation", inference = "bootstrap", nboot = 1000)
```

```{r}
summary(res_gformula.m.a)
```

# {-}

## Imai Mediation Approach
Using the mediation package approach; not used

Mediator model (PC3 in this case)
```{r}
med.fit <- lm(Dim.3 ~ phenotype + sex + ga + batch, data = df)
print(tidy(med.fit))
```
Interestingly, GA is only marginally associated with FLT1 status without cell composition in the model
```{r}
out.fit <- lm(FLT1 ~ phenotype + sex + ga + batch, data = df)
print(tidy(out.fit))
```

Outcome model
```{r}
out.fit <- lm(FLT1 ~ Dim.3 + phenotype + sex + ga + batch, data = df)
print(tidy(out.fit))
```

```{r}
# Bayesian quasi-CIs
#med.out <- mediate(med.fit, out.fit, treat = "phenotype", mediator = "Dim.3", robustSE = TRUE, sims = 100)
#summary(med.out)

# Bootstrap CIs
med.out.boot <- mediation::mediate(med.fit, out.fit, treat = "phenotype", mediator = "Dim.3", boot = TRUE, boot.ci.type = "bca", sims = 2000)
summary(med.out.boot)
```

```{r}
model.m  <- lm(Dim.3 ~ phenotype + sex + ga + batch, data = df)
print(tidy(med.fit))

# Interaction model to allow heterogeneity between the two treatment levels
model.y <- lm(FLT1 ~ Dim.3*phenotype + sex + ga + batch, data = df)
print(tidy(out.fit))

med.out.boot <- mediation::mediate(model.m, model.y, treat = "phenotype", mediator = "Dim.3", boot = TRUE, boot.ci.type = "bca", sims = 2000)
summary(med.out.boot)

# Interaction is not statistically significant
test.TMint(med.out.boot, conf.level = 0.95)
```

```{r}
# Convert phenotype to treatment dummy variable for mediation modelling
df$treat <- ifelse(df$phenotype == "Control", 0, 1)

# Pretreatment covariates
Xnames <- c("sex", "batch")
m.med <- multimed(outcome = "FLT1", med.main = "Dim.3", med.alt = "ga",
                  treat = "treat", covariates = Xnames,
                  data = df, sims = 2000)
summary(m.med)
```

```{r}
plot(m.med, type = "point")
```

```{r}
plot(m.med, type = c("sigma", "R2-total"), tgroup = c("treated", "control"))
```
