---
title: "Validation of scRNA-seq deconvolution approach"
author: "Kyle Campbell"
date: "3/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

#install.packages("package_name",  repos="http://cran.us.r-project.org") # Only this mirror worked for multiple packages
library(devtools)
library(here)
library(ggrepel)
options(ggrepel.max.overlaps = Inf) # increase max.overlaps for ggrepel
library(ggpubr)
library(gprofiler2)
library(ggridges)
library(knitr)
library(tidyverse)
library(Seurat)
#remotes::install_github('satijalab/seurat-wrappers')
library(SeuratWrappers)
library(cowplot)
library(rgl)
library(EnvStats)
library(reticulate)
library(clustree)
library(Matrix)

# Bioconductor packages
#remotes::install_github('chris-mcginnis-ucsf/DoubletFinder') #Install DoubletFinder and depends
library(batchelor)
library(edgeR)
library(scater)
# Set seed since this pipeline includes stochastic algorithms
set.seed(1)
```

## Custom functions

Take average rmse between two vectors
```{r}
rmse <- function(actual, predicted) {
  return(sqrt(mean((actual - predicted)^2)))
}
```

Function to format ~/CIBERSORT_Adjusted.txt CIBERSORTx output
```{r}
get_imputed_matrix <- function(res) {
  imp <- res %>%
  dplyr::select(!c("P-value", "Correlation", "RMSE")) %>%
  t() %>%
  as.data.frame()
imp <- imp %>% rownames_to_column("cell.type")
return(imp)
}
```

Function to get cell type proportions from Seurat object
```{r}
# https://stackoverflow.com/questions/26003574/use-dynamic-name-for-new-column-variable-in-dplyr
# glue syntax to programmatically create name in dplyr function
get_seu_cell_percentages <- function(seu, export_name) {
  x <-
    seu@meta.data %>%
    group_by(cell.type, .drop = F) %>%
    tally() %>%
    dplyr::mutate("{export_name}" := (n / sum(n))) %>%
    dplyr::select(cell.type, all_of(export_name))
  return(x)
}
```

Function to get cell type proportions by biological replicate
```{r}
get_seu_cell_percentages_by_sample <- function(seu, export_name) {
  x <-
    seu@meta.data %>%
    group_by(biorep, cell.type, .drop = F) %>%
    summarise(cnt = n(), group = "keep") %>%
    dplyr::mutate(freq = round(cnt/sum(cnt), 3)) %>%
    dplyr::select(biorep, cell.type, all_of(export_name))
  return(x)
}
```


## Load data

Load deconvoluted microarray data and drop preeclampsia samples
```{r}
pheno <- readRDS(here("data", "GSE75010_with_abundances_2022-05-01.rda"))
controls <- pheno %>%
  filter(phenotype == "C") %>%
  pull(sample)
```


```{r}
ma <- read_tsv(here("results", "cibersortx_local", "analytic_gse75010", "CIBERSORTx_Adjusted.txt"))
ma$Source <- "GSE75010"

ma <- ma %>%
  dplyr::select(!c(RMSE, Correlation, `P-value`)) %>%
  filter(Mixture %in% controls) %>%
  dplyr::select(Mixture, Source, everything()) %>%
  pivot_longer(cols = !c(Mixture, Source), names_to = "cell.type", values_to = "freq") %>%
  mutate(freq = round(freq, 3))
```

Verify each subgroup adds to 100%
```{r}
ma %>% group_by(Mixture) %>% summarise(total = sum(freq)) %>% summarise(total.total = sum(total))
```

Load deconvoluted sorted cell type data (to pull whole villous tissue samples)
```{r}
sort <- read_tsv(here("results", "cibersortx_local", "analytic_kc_facs_mixtures", "CIBERSORTx_Adjusted.txt"))
sort$Source <- "Sorted"
sort <- sort %>%
  filter(grepl("Composite", x = Mixture))
sort[2, 1] <- "Sorted_1A_Composite"
sort[5, 1] <- "Sorted_1B_Composite"
sort <- sort %>%
  dplyr::select(!c(RMSE, Correlation, `P-value`)) %>%
  dplyr::select(Mixture, Source, everything()) %>%
  pivot_longer(cols = !c(Mixture, Source), names_to = "cell.type", values_to = "freq") %>%
  mutate(freq = round(freq, 3))
```

Verify each subgroup adds to 100%
```{r}
sort %>% group_by(Mixture) %>% summarise(total = sum(freq))
```

Load cleaned Seurat data
```{r}
#seu <- readRDS(here("data", "analytic", "2022-03-15_all_merged_pruned_batch_mnn.rda"))
seu <- readRDS(here("data", "analytic", "2022-05-01_all_merged_pruned_batch_mnn.rda"))
seu$cell.type <- factor(seu$cell.type)
```

## Check tissue cell type proportions

Get cell type proportions from sc data
```{r}
sc <- get_seu_cell_percentages_by_sample(seu, "freq")
sc$Source <- "Single-cell"
colnames(sc)[1] <- "Mixture"
```

Check that each subgroup adds to 100%
```{r}
sc %>% summarise(total = sum(freq))
```

```{r}
merge <- rbind(sc, sort, ma)
```

Take averages by Source for each cell type for correlation comparison
```{r}
merge.summary <- 
  merge %>%
  group_by(cell.type, Source) %>%
  summarise(freq.average = mean(freq))

merge.summary.wide <- merge.summary %>%
  pivot_wider(names_from = `cell.type`, values_from = `freq.average`) %>%
  column_to_rownames(var = "Source") %>%
  t() %>%
  as.data.frame()
```

Check freq.averages is about ~100%
```{r}
colSums(merge.summary.wide)
dim(merge.summary.wide)
```

```{r}
#stats::cor.test(x = merge.summary.wide$GSE75010, y = merge.summary.wide$Sorted)
```

```{r}
merge.fetal <- merge %>%
  dplyr::filter(grepl("Fetal ", cell.type))

merge.maternal <- merge %>%
  dplyr::filter(grepl("Maternal ", cell.type))
```


c("GSE75010", "Single-cell", "Sorted")
```{r}
ggplot(merge.fetal, aes(x = freq, y = cell.type, color = Source, point_color = Source, fill = Source)) +
  geom_density_ridges(
    jittered_points = TRUE, scale = .95, rel_min_height = .01,
    point_shape = "|", point_size = 1, size = 0.25,
    position = position_points_jitter(height = 0),
    alpha = 0.25
  ) +
  ylab("Cell Type") +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0), name = "Cell Type Proportion", labels = scales::percent_format(accuracy = 1)) +
  coord_cartesian(clip = "off") +
  ggtitle("Whole tissue cell type proportions - fetal cell types") +
  theme_ridges(center = TRUE) +
  theme(legend.title = element_blank(), legend.position = c(.75, .85), legend.text = element_text(size = 10), axis.title.x=element_blank(), plot.title = element_text(size = 12), axis.text.y = element_text(size = 8), axis.title.y = element_blank())

fetal.panel <- ggplot(merge.fetal, aes(x = freq, y = cell.type, color = Source, point_color = Source, fill = Source)) +
  geom_density_ridges(
    jittered_points = TRUE, scale = .95, rel_min_height = .01,
    point_shape = "|", point_size = 1, size = 0.25,
    position = position_points_jitter(height = 0),
    alpha = 0.25
  ) +
  ylab("Cell Type") +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0), name = "Cell Type Proportion", labels = scales::percent_format(accuracy = 1)) +
  coord_cartesian(clip = "off") +
  ggtitle("Whole tissue cell type proportions - fetal cell types") +
 theme_minimal(base_size = 20) +
  theme(axis.text.y = element_text(color = "black"))
```

```{r}
ggplot(merge.maternal, aes(x = freq, y = cell.type, color = Source, point_color = Source, fill = Source)) +
  geom_density_ridges(
    jittered_points = TRUE, scale = .95, rel_min_height = .01,
    point_shape = "|", point_size = 1, size = 0.25,
    position = position_points_jitter(height = 0),
    alpha = 0.25
  ) +
  ylab("Cell Type") +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0), name = "Cell Type Proportion", labels = scales::percent_format(accuracy = 1)) +
  coord_cartesian(clip = "off") +
  ggtitle("Whole tissue cell type proportions - maternal cell types") +
  theme_ridges(center = TRUE) +
  theme(legend.title = element_blank(), legend.position = c(.75, .85), legend.text = element_text(size = 10), axis.title.x=element_blank(), plot.title = element_text(size = 12), axis.text.y = element_text(size = 8), axis.title.y = element_blank())

maternal.panel <- ggplot(merge.maternal, aes(x = freq, y = cell.type, color = Source, point_color = Source, fill = Source)) +
  geom_density_ridges(
    jittered_points = TRUE, scale = .95, rel_min_height = .01,
    point_shape = "|", point_size = 1, size = 0.25,
    position = position_points_jitter(height = 0),
    alpha = 0.25
  ) +
  ylab("") +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0), name = "Cell Type Proportion", labels = scales::percent_format(accuracy = 1)) +
  coord_cartesian(clip = "off") +
  ggtitle("Whole tissue cell type proportions - maternal cell types") +
 theme_minimal(base_size = 20) +
  theme(axis.text.y = element_text(color = "black"))
```

```{r}
prop.panel <- ggarrange(fetal.panel, maternal.panel, labels = "AUTO", common.legend = TRUE, legend = "bottom", font.label = list(size = 24, face = "bold", color = "black"))
#ggexport(prop.panel, width = 1920, height = 1080, filename = here("results", "deconvolution_validation", paste0(Sys.Date(), "_fetal_maternal_proportions.png")))
```


## Training/test split

OUTMODED - Split 50/50 test/training dataset, randomly. Leads to significant differences due to large df
```{r, eval = F, include = F}
# OUTMODED
set.seed(1)
fraction <- 0.5
n.cell <- nrow(seu@meta.data)
n.cell.subsample <- fraction*n.cell
subsample.train <- sample(colnames(seu), n.cell.subsample, replace = FALSE)

seu.train <- subset(seu, cells = subsample.train)
seu.test <- subset(seu, cells = subsample.train, invert = T)
```

Split 50/50 forcing equal proportions by cell type; when cell type count is odd, the extra cell goes to training
```{r, eval = F}
set.seed(1)
# Split seu@meta.data into cell.type groups
x.groups <-seu@meta.data %>%
  rownames_to_column(var = "cell.id") %>%
  dplyr::select(cell.id, cell.type) %>%
  group_by(cell.type) %>%
  group_split()
# Sample at 50% from those groups w/o replacement
cell.ids.sample <- x.groups %>% map(~ sample(x = .x$cell.id, size = ceiling(nrow(.x)/2), replace = F)) %>% unlist

seu.train <- subset(seu, cells = cell.ids.sample)
seu.test <- subset(seu, cells = cell.ids.sample, invert = T)

# Save training IDs for reproducibility
#saveRDS(cell.ids.sample, here("data", paste0(Sys.Date(), "deconvolution_validation_training_cell_ids.Rda")))
```

Load data
```{r}
cell.ids.sample <- readRDS(here("data", "2022-05-10deconvolution_validation_training_cell_ids.Rda"))
seu.train <- subset(seu, cells = cell.ids.sample)
seu.test <- subset(seu, cells = cell.ids.sample, invert = T)
```


UMAP plots of training and test data are visually identical
```{r}
DimPlot(seu.train, pt.size = 0.25, label = T, repel = T) + NoLegend() + ggtitle("Training")
DimPlot(seu.train, pt.size = 0.25, label = T, repel = T) + NoLegend() + ggtitle("Test")
```


```{r}
table(seu.train$sex, seu.train$fetal)
table(seu.test$sex, seu.test$fetal)
```

```{r}
all.cell.types <- seu$cell.type %>% factor %>% levels
train.celltypes <- seu.train@meta.data %>% group_by(cell.type) %>% tally()
colnames(train.celltypes)[2] <- "train"

test.celltypes <- seu.test@meta.data %>% group_by(cell.type) %>% tally()
colnames(test.celltypes)[2] <- "test"

cell.type.counts <- left_join(train.celltypes, test.celltypes) %>%
  mutate(test5050.proportions = test/sum(test))
```

```{r}
# Human-readable percentage formatting
#mutate("Percentage" = round(test/sum(test)*100, digits = 2))
```

No significant difference in cell type counts between train and test datasets via Pearson correlation test and rmse
```{r}
cor.test(x = cell.type.counts$train, y = cell.type.counts$test)
rmse(actual =  cell.type.counts$test, predicted = cell.type.counts$train) # note on absolute, not percent scale
```

Save seu object of training cell ids
```{r}
#saveRDS(seu.train, here("data", "seurat", paste0(Sys.Date(), "_kc_pr_ts_5050_train_sig_input.rda")))
```   

Format test data for export to CibersortX; start with the whole training sample, 100% fetal and 100% maternal
```{r}
format_seu_for_cibersortx_mixture <- function(seu, export_name) {
  pseudobulk <- seu@assays$RNA %>% rowSums() %>% cpm() %>% as.data.frame() %>% rownames_to_column(var = "GeneSymbol")
  colnames(pseudobulk)[2] <- export_name
  return(pseudobulk)
}

test.pseudobulk <- format_seu_for_cibersortx_mixture(seu = seu.test, export_name = "Test_5050")
test.fetal <- format_seu_for_cibersortx_mixture(seu = subset(seu.test, subset = fetal == "Fetal"), export_name = "Fetal_5050")
test.maternal <- format_seu_for_cibersortx_mixture(seu = subset(seu.test, subset = fetal == "Maternal"), export_name = "Maternal_5050")
```

## Testing fetal sex deconvolution
Baseline distribution is 85% fetal
Of fetal, 41% are female (7,110 cells)
1,284 maternal cells will make up about 15% of 8,561 total cells; randomly select those maternal cells once and mix with all fetal or all male cells (downsample male to 7277 randomly selected cells)
We have 3,106 total maternal cells in the dataset

```{r}
set.seed(1)
# Split seu.test into sex,fetal groups for three unique groups maternal female, fetal male, and fetal female
x.groups <-seu.test@meta.data %>%
  rownames_to_column(var = "cell.id") %>%
  dplyr::select(cell.id, cell.type, fetal, sex) %>%
  group_by(sex) %>%
  group_by(fetal, .add = T) %>%
  group_split()
names(x.groups) <- c("female.fetal", "female.maternal", "male.fetal") 

sex.test.maternal.ids <- sample(x.groups$female.maternal$cell.id, size = 1284, replace = F)
sex.test.female.fetal.ids <- x.groups$female.fetal$cell.id
sex.test.male.fetal.ids <- sample(x.groups$male.fetal$cell.id, size = 7110, replace = F)

testing.ids <- list(sex.test.maternal.ids, sex.test.female.fetal.ids, sex.test.male.fetal.ids)
#saveRDS(testing.ids, here("data", paste0(Sys.Date(), "deconvolution_sex_testing_ids.rda"))) # saved 5/10/2022; replaced with 7/21/2022 due to down-sample size typo; now fixed and used in final analysis
```

Load data
```{r}
#testing.ids <- readRDS(here("data", "2022-05-21deconvolution_sex_testing_ids.rda"))
testing.ids <- readRDS(here("data", "2022-07-21deconvolution_sex_testing_ids.rda"))
```

```{r}
sex.test.female <- format_seu_for_cibersortx_mixture(subset(seu.test, cells = c(sex.test.female.fetal.ids, sex.test.maternal.ids)), export_name = "sex.test.female")

sex.test.male <- format_seu_for_cibersortx_mixture(subset(seu.test, cells = c(sex.test.male.fetal.ids, sex.test.maternal.ids)), export_name= "sex.test.male")
```

## Join pseudobulk mixtures and export for CibersortX deconvolution 

```{r}
test <- test.pseudobulk %>%
  left_join(test.fetal, by = "GeneSymbol") %>%
  left_join(test.maternal, by = "GeneSymbol") %>%
  left_join(sex.test.female, by = "GeneSymbol") %>%
  left_join(sex.test.male, by = "GeneSymbol")
```

Write 
```{r}
#write.table(x=test, sep = '\t', quote = F, row.names = F, col.names = T, file = here("data", "cibersortx_local", paste0("kc_pr_ts_sc_5050_test_mixtures_cpm_input_", Sys.Date(), ".txt")))
```

Possibly cpm() the sc-sig-matrix input? CibersortX online tutorial 1 claims it does this automatically, but the file input page does not say that
```{r}
#read.table(kc_pr_ts_sc_5050_train_signature_matrix_input_2022-03-25.txt
```

## Get test pseudobulk mixture percentages

Function to get actual cell type percentages
```{r}
Test_5050_actual <- get_seu_cell_percentages(seu = seu.test, export_name = "Test_5050_actual")
Fetal_5050_actual <- get_seu_cell_percentages(subset(seu.test, subset = fetal == "Fetal"), export_name = "Fetal_5050_actual")
Maternal_5050_actual <- get_seu_cell_percentages(subset(seu.test, subset = fetal == "Maternal"), export_name = "Maternal_5050_actual")
Female_SexTest_actual <- get_seu_cell_percentages(subset(seu.test, cells = c(sex.test.female.fetal.ids, sex.test.maternal.ids)), export_name = "sex.test.female_actual")
Male_SexTest_actual <- get_seu_cell_percentages(subset(seu.test, cells = c(sex.test.male.fetal.ids, sex.test.maternal.ids)), export_name = "sex.test.male_actual")
# Join the results
#Test <- left_join(Test_5050_actual, left_join(Fetal_5050_actual, Maternal_5050_actual))
Test <- left_join(Test_5050_actual, Fetal_5050_actual) %>%
  left_join(Maternal_5050_actual) %>%
  left_join(Female_SexTest_actual) %>%
  left_join(Male_SexTest_actual)
```

## Load CibersortX results
Verify deconvolution results
```{r}
res.S <- read_tsv(file = here("results", "cibersortx_local", "pseudobulk_testing_gmax2000_perm_50_S", "CIBERSORTx_Adjusted.txt")) %>%
  column_to_rownames("Mixture")
res.noS <- read_tsv(file = here("results", "cibersortx_local", "test_train_5050_mixtures_gmax2000_perm_50_noS", "CIBERSORTX_Results.txt")) %>%
  column_to_rownames("Mixture")
```

Function to drop CibersortX deconvolution statistics and reformat with cell type as a column 

```{r}
res.df.S <- get_imputed_matrix(res.S)
res.df.noS <- get_imputed_matrix(res.noS)
```

## S-mode batch correction and fetal maternal test data with g.max = 2000 {.tabset}

```{r}
ggplot.dat <- left_join(Test, res.df.S)
```

### All cell types at natural rates
Batch correction significantly overperforms in full test set
```{r}
cor.test(x = res.df.S$Test_5050, y = Test$Test_5050_actual)
rmse(actual = res.df.S$Test_5050, predicted = Test$Test_5050_actual)
plot(x = res.df.S$Test_5050, y = Test$Test_5050_actual)
cor.test(x = res.df.noS$Test_5050, y = Test$Test_5050_actual)
rmse(actual = res.df.noS$Test_5050, predicted = Test$Test_5050_actual)
plot(x = res.df.noS$Test_5050, y = Test$Test_5050_actual)
```

### Test data includes only fetal cells
Batch correction significantly overperforms in fetal test set
```{r}
cor.test(x = res.df.S$Fetal_5050, y = Test$Fetal_5050_actual)
rmse(predicted = res.df.S$Fetal_5050, actual = Test$Fetal_5050_actual)
plot(x = res.df.S$Fetal_5050, y = Test$Fetal_5050_actual)
cor.test(x = res.df.noS$Fetal_5050, y = Test$Fetal_5050_actual)
rmse(predicted = res.df.noS$Fetal_5050, actual = Test$Fetal_5050_actual)
plot(x = res.df.noS$Fetal_5050, y = Test$Fetal_5050_actual)
```

### Test data includes only maternal cells
No difference in maternal test set; slight advantage in effect measure for no batch correction
```{r}
cor.test(x = res.df.S$Maternal_5050, y = Test$Maternal_5050_actual)
rmse(predicted = res.df.S$Maternal_5050, actual = Test$Maternal_5050_actual)
plot(x = res.df.S$Maternal_5050, y = Test$Maternal_5050_actual)
cor.test(x = res.df.noS$Maternal_5050, y = Test$Maternal_5050_actual)
rmse(predicted = res.df.noS$Maternal_5050, actual = Test$Maternal_5050_actual)
plot(x = res.df.noS$Maternal_5050, y = Test$Maternal_5050_actual)
```

### Test data only female fetal cells
Deconvolution performs just as well in only female fetal cells
```{r}
cor.test(x = res.df.S$sex.test.female, y = Test$sex.test.female_actual)
rmse(predicted = res.df.S$sex.test.female, actual = Test$sex.test.female_actual)
plot(x = res.df.S$sex.test.female, y = Test$sex.test.female_actual)
```

### Test data only male fetal cells
Deconvolution does not perform as well in only male fetal cells; the correlation 
```{r}
cor.test(x = res.df.S$sex.test.male, y = Test$sex.test.male_actual)
rmse(predicted = res.df.S$sex.test.male, actual = Test$sex.test.male_actual)
plot(x = res.df.S$sex.test.male, y = Test$sex.test.male_actual)

ggplot(data = ggplot.dat,
       aes(x = sex.test.male,
           y = sex.test.male_actual,
           label = ifelse(
            abs(sex.test.male_actual - sex.test.male) > .03,
             cell.type,
             ""
           ))) +
  geom_point() +
  geom_text_repel(min.segment.length = 0)
```

## {-}

## Analytic test data w/ default g-max (300min to 500max) {.tabset}
Verify deconvolution results
```{r}
#res.S <- read_tsv(file = here("results", "cibersortx_local", "testing_gmax_default_perm_50_S", "CIBERSORTx_Adjusted.txt")) %>%
#  column_to_rownames("Mixture")

# Updated analysis
res.S <- read_tsv(file = here("results", "cibersortx_local", "analytic_deconvolution_validation_testing", "CIBERSORTx_Adjusted.txt")) %>%
  column_to_rownames("Mixture")

# Reformat for analysis
res.df.S <- get_imputed_matrix(res.S)

base.size <- 8
```

### All cell types at natural rates

```{r}
predicted = res.df.S$Test_5050
actual = Test$Test_5050_actual

gg.data <- data.frame(
  predicted = predicted,
  actual = actual
)

cor.test(x = actual, y = predicted)
pearson <- round(cor(x = actual, y = predicted), 3)
pearson
rmse.deconvolution = round(rmse(actual = actual, predicted = predicted), 3)
rmse.deconvolution

plot <- ggplot(gg.data, aes(x = predicted, y = actual)) + geom_point() + labs(subtitle = paste0("Correlation = ", pearson, ", Root mean square error = ", rmse.deconvolution)) + xlab("Predicted cell type proportions") + ylab("Actual cell type proportions") + ggtitle("Deconvolution validation - 50/50 split") + theme_bw(base_size = base.size) + geom_smooth(method='lm', formula= y~x)

# Store plot for gg panel later
plot.test.5050 <- plot

plot

plot(x = res.df.S$Test_5050, y = Test$Test_5050_actual)
```

### Test data includes only fetal cells

```{r}
predicted = res.df.S$Fetal_5050
actual = Test$Fetal_5050_actual

gg.data <- data.frame(
  predicted = predicted,
  actual = actual
)

cor.test(x = actual, y = predicted)
pearson <- round(cor(x = actual, y = predicted), 3)
pearson
rmse.deconvolution = round(rmse(actual = actual, predicted = predicted), 3)
rmse.deconvolution

plot <- ggplot(gg.data, aes(x = predicted, y = actual)) + geom_point() + labs(subtitle = paste0("Correlation = ", pearson, ", Root mean square error = ", rmse.deconvolution)) + xlab("Predicted cell type proportions") + ylab("Actual cell type proportions") + ggtitle("Only fetal cells present") + theme_bw(base_size = base.size) + geom_smooth(method='lm', formula= y~x)

# Store plot for gg panel later
plot.fetal.5050 <- plot

plot

plot(x = res.df.S$Fetal_5050, y = Test$Fetal_5050_actual)
```

### Test data includes only maternal cells

```{r}
predicted = res.df.S$Maternal_5050
actual = Test$Maternal_5050_actual

gg.data <- data.frame(
  predicted = predicted,
  actual = actual
)

cor.test(x = actual, y = predicted)
pearson <- round(cor(x = actual, y = predicted), 3)
pearson
rmse.deconvolution = round(rmse(actual = actual, predicted = predicted), 3)
rmse.deconvolution

plot <- ggplot(gg.data, aes(x = predicted, y = actual)) + geom_point() + labs(subtitle = paste0("Correlation = ", pearson, ", Root mean square error = ", rmse.deconvolution)) + xlab("Predicted cell type proportions") + ylab("Actual cell type proportions") + ggtitle("Only maternal cells present") + theme_bw(base_size = base.size) + geom_smooth(method='lm', formula= y~x) #+ theme(plot.subtitle=element_text(size=base.size*.5))

# Store plot for gg panel later
plot.maternal.5050 <- plot

plot

plot(x = res.df.S$Maternal_5050, y = Test$Maternal_5050_actual)
```

### Test data only female fetal cells

```{r}
predicted = res.df.S$sex.test.female
actual = Test$sex.test.female_actual

gg.data <- data.frame(
  predicted = res.df.S$Test_5050,
  actual = Test$Test_5050_actual
)

cor.test(x = actual, y = predicted)
pearson <- round(cor(x = actual, y = predicted), 3)
pearson
rmse.deconvolution = round(rmse(actual = actual, predicted = predicted), 3)
rmse.deconvolution

plot <- ggplot(gg.data, aes(x = predicted, y = actual)) + geom_point() + labs(subtitle = paste0("Correlation = ", pearson, ", Root mean square error = ", rmse.deconvolution)) + xlab("Predicted cell type proportions") + ylab("Actual cell type proportions") + ggtitle("Only female fetal cells present") + theme_bw(base_size = base.size) + geom_smooth(method='lm', formula= y~x)

# Store plot for gg panel later
plot.sex.test.female <- plot

plot

plot(x = res.df.S$sex.test.female, y = Test$sex.test.female_actual)
```

### Test data only male fetal cells

```{r}
predicted = res.df.S$sex.test.male
actual = Test$sex.test.male_actual

gg.data <- data.frame(
  predicted = res.df.S$Test_5050,
  actual = Test$Test_5050_actual
)

cor.test(x = actual, y = predicted)
pearson <- round(cor(x = actual, y = predicted), 3)
pearson
rmse.deconvolution = round(rmse(actual = actual, predicted = predicted), 3)
rmse.deconvolution

plot <- ggplot(gg.data, aes(x = predicted, y = actual)) + geom_point() + labs(subtitle = paste0("Correlation = ", pearson, ", Root mean square error = ", rmse.deconvolution)) + xlab("Predicted cell type proportions") + ylab("Actual cell type proportions") + ggtitle("Only male fetal cells present") + theme_bw(base_size = base.size, ) + geom_smooth(method='lm', formula= y~x)

# Store plot for gg panel later
plot.sex.test.male <- plot

plot

plot(x = res.df.S$sex.test.male, y = Test$sex.test.male_actual)
```

## { - }

Optimal file export settings are much different from display in R
```{r}
panel <- ggarrange(plot.test.5050, plot.fetal.5050, plot.maternal.5050, plot.sex.test.female, plot.sex.test.male, ncol = 3, nrow = 2, labels = "auto", font.label = list(size = 16, face = "bold", color = "black"))
print(panel)
# Works okay with theme_bw(base_size = 20) and font.label size = 24 in ggarrange
#ggexport(panel, filename = here("results", "deconvolution_validation", paste0(Sys.Date(), "_deconvolution_validation_multipanel.png")), width = 1920, height = 1080)
```

Need to run the png() dev.off() command in the console to work
```{r, eval = F}
png(filename = here("results", "deconvolution_validation", paste0(Sys.Date(), "_deconvolution_validation_multipanel.png")), width = 2816, height = 2160, units = "px", res = 300)
panel
dev.off()
```

## test data w/ g-max = 1000 {.tabset}

Verify deconvolution results
```{r, eval = F, include = F}
res.S <- read_tsv(file = here("results", "cibersortx_local", "testing_gmax_1000_perm_50_S", "CIBERSORTx_Adjusted.txt")) %>%
  column_to_rownames("Mixture")
res.df.S <- get_imputed_matrix(res.S)
```

### All cell types at natural rates

```{r, eval = F, include = F}
cor.test(x = res.df.S$Test_5050, y = Test$Test_5050_actual)
rmse(actual = res.df.S$Test_5050, predicted = Test$Test_5050_actual)
plot(x = res.df.S$Test_5050, y = Test$Test_5050_actual)
```

### Test data includes only fetal cells

```{r, eval = F, include = F}
cor.test(x = res.df.S$Fetal_5050, y = Test$Fetal_5050_actual)
rmse(predicted = res.df.S$Fetal_5050, actual = Test$Fetal_5050_actual)
plot(x = res.df.S$Fetal_5050, y = Test$Fetal_5050_actual)
```

### Test data includes only maternal cells

```{r, eval = F, include = F}
cor.test(x = res.df.S$Maternal_5050, y = Test$Maternal_5050_actual)
rmse(predicted = res.df.S$Maternal_5050, actual = Test$Maternal_5050_actual)
plot(x = res.df.S$Maternal_5050, y = Test$Maternal_5050_actual)
```

### Test data only female fetal cells

```{r, eval = F, include = F}
cor.test(x = res.df.S$sex.test.female, y = Test$sex.test.female_actual)
rmse(predicted = res.df.S$sex.test.female, actual = Test$sex.test.female_actual)
plot(x = res.df.S$sex.test.female, y = Test$sex.test.female_actual)
```

### Test data only male fetal cells

```{r, eval = F, include = F}
cor.test(x = res.df.S$sex.test.male, y = Test$sex.test.male_actual)
rmse(predicted = res.df.S$sex.test.male, actual = Test$sex.test.male_actual)
plot(x = res.df.S$sex.test.male, y = Test$sex.test.male_actual)
```

## { - }