---
title: "facs_sort_deconvolution"
author: "Kyle Abraham Campbell"
date: "4/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(factoextra)
library(tidyverse)
library(betareg)
library(boot)
library(psych)
library(readxl)
```

Import CIBERSORT cell type fraction imputation data.
```{r}
frac <- read_tsv(here("results", "cibersortx_local", "analytic_kc_facs_mixtures", "CIBERSORTx_Adjusted.txt"))
colnames(frac)[1] <- "sample"
head(frac)
dim(frac)
```

```{r}
levels(factor(frac$sample))
colnames(frac[2:28])
```

Summary stats
```{r}
describe(frac$`P-value`)
describe(frac$Correlation)
describe(frac$RMSE)
```

```{r}
describe(frac$`Fetal Cytotrophoblasts`)
describe(frac$`Fetal Syncytiotrophoblast`)
describe(frac$`Fetal Naive CD8+ T Cells`)
```

```{r}
describe(frac$`Maternal Naive CD8+ T Cells`)
describe(frac$`Maternal Natural Killer Cells`)
describe(frac$`Maternal Plasma Cells`)
```

```{r}
facs.cats <-
  frac %>%
  mutate(Cytotrophoblast = round(rowSums(across((matches("Cytotrophoblast")))), 2)) %>%
  mutate(Fibroblast = round(rowSums(across((matches("Fibroblast|Mesenchymal")))), 2)) %>%
  mutate(Leukocyte = round(rowSums(across((matches("T Cells|Monocytes|Natural Killer|B Cells|Dendritic Cells|Plasma Cells")))), 2)) %>%
  mutate(cell.type = substr(sample, start = 10, stop = 99)) %>%
  mutate(`Fetal Endothelial Cells` = round(`Fetal Endothelial Cells`, 2)) %>%
  mutate(`Fetal Extravillous Trophoblasts` = round(`Fetal Extravillous Trophoblasts`, 2)) %>%
  mutate(`Fetal Hofbauer Cells` = round(`Fetal Hofbauer Cells`, 2)) %>%
  mutate(`Fetal Syncytiotrophoblast` = round(`Fetal Syncytiotrophoblast`, 2)) %>%
  mutate(`P-value` = round(`P-value`, 2)) %>%
  mutate(`Correlation` = round(`Correlation`, 2)) %>%
  mutate(`RMSE` = round(`RMSE`, 2)) %>%
  dplyr::select(sample, cell.type, Cytotrophoblast, `Fetal Endothelial Cells`, `Fetal Extravillous Trophoblasts`, `Fetal Hofbauer Cells`, Fibroblast, Leukocyte, `Fetal Syncytiotrophoblast`) %>%
  arrange(cell.type)
colnames(facs.cats)[1] <- "Sample"
colnames(facs.cats)[2] <- "Cell Type"
```

```{r}
list <- list(
  "Deconvoluted sorted cell type" = facs.cats,
  "Deconvolution Results" = frac
  )

write.xlsx(list, file = here("results", "bulk", paste0(Sys.Date(), "_sorted_deconvolution_res.xlsx")))
#write_csv(facs.cats, file = here("data", paste0(Sys.Date(), "_facs_mixtures_deconvoluted.csv")))
```

Notes for 05-01-2022 results, which had apparent sample swap:

The 105xxx samples look alright-ish, can make a reasonable argument for enrichment on most of them (but still rough), though Syncytiotrophoblast was still predicted at 0% for that sample

Sample 7 looks fine for the most part (KC39), failure to distinguish between Hofbauer and leukocyte, same issue with syncytiotrophoblast

Looks like Sample 6 may have a LK/HB swap, which is reasonable considering they are sorted right next to each other
Have the same LK/HB swap in Sample 1 

Next step is to check the transcriptomic data to see if reasonable.
Can also check FACS-sorting records to see if Samples 6 and 1 were sorted or RNA extractd on the same day whereas the other 2 samples were not

Cite dissociation and freezing bias issues
Cite previous paper that describes CT as being unamenable to sorting; scRNA-seq results are much stronger 

```{r}
fb <- facs.cats %>% filter(`Cell Type` == "Fibroblast") %>% pull(Fibroblast)
mean(fb)
sd(fb)
```

```{r}
lk <- facs.cats %>% filter(`Cell Type` == "Leukocyte") %>% pull(Leukocyte)
mean(lk)
sd(lk)
```
```{r}
hb <- facs.cats %>% filter(`Cell Type` == "Hofbauer") %>% pull(Leukocyte)
mean(hb)
sd(hb)
```

