---
title: "KC and PR scRNA-seq import of filtered 10x output and QC of placental villous tissue"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

#install.packages("package_name",  repos="http://cran.us.r-project.org") # Only this mirror worked for multiple packages
library(knitr)
library(tidyverse)
library(devtools)
library(Seurat)
#remotes::install_github('satijalab/seurat-wrappers')
library(SeuratWrappers)
library(cowplot)
library(rgl)
library(EnvStats)
library(reticulate)
library(clustree)
library(Matrix)

#remotes::install_github('chris-mcginnis-ucsf/DoubletFinder') #Install DoubletFinder and depends
library(DoubletFinder)

# Bioconductor packages
library(DropletUtils)
library(edgeR)
library(scater)
# Set seed since this pipeline includes stochastic algorithms
set.seed(1)
# Set default working directory
data_dir <- "C:/Users/k_cam/Google Drive/Placenta_Cell_Types/RNA/placenta_rna/data/"
results_dir <- "C:/Users/k_cam/Google Drive/Placenta_Cell_Types/RNA/placenta_rna/results/single_cell_qc/"
```
### Important References
Minimal QC recommended: Current best practices in single‐cell RNA‐seq analysis: a tutorial
Orchestrating Single-cell Analysis
https://bioconductor.org/packages/release/bioc/vignettes/DropletUtils/inst/doc/DropletUtils.html
### Compiling Issues
This document had trouble compiling via knitr and pandoc. The following thread fixed the issue (https://github.com/rstudio/rstudio/issues/3661).
## Custom function to create list of Seurat objects from 10x output directories
```{r}
# Function to convert from 10x output to seurat object with no filtering
tenx_to_seu <- function(tenx_directory, name) {
  seu <- CreateSeuratObject(
    counts = Read10X(tenx_directory),
    project = name)
  return(seu)
}
```
## Get list of Seurat objects
```{r}
raw_dir <- paste0(data_dir, "tenx_outs/")
# Pull full file paths
names <- 
  grep(
    dir(
      path = raw_dir             #list all files, unfortunately includes .ini files
      ),
    pattern = "(kc|pr)(.\\d+)+", #pattern to search for, excludes .ini files
    value = T                    #return matches
    )
files <-
  paste0(
    raw_dir, 
    names
  )
# Create list of Seurat objects
counts <- map2(files, names, tenx_to_seu)
# Name list
names(counts) <- names
# Manually add batch metadata to distinguish kc from pr
counts$kc.40.1$'batch' <- "kc"
counts$kc.40.2$'batch' <- "kc"
counts$kc.42.1$'batch' <- "kc"
counts$kc.42.2$'batch' <- "kc"
counts$pr.478$'batch' <- "pr"
counts$pr.481$'batch' <- "pr"
counts$pr.484$'batch' <- "pr"
# Manually add bioreplicate metadata
counts$kc.40.1$'biorep' <- "kc.40"
counts$kc.40.2$'biorep' <- "kc.40"
counts$kc.42.1$'biorep' <- "kc.42"
counts$kc.42.2$'biorep' <- "kc.42"
counts$pr.478$'biorep' <- "pr.478"
counts$pr.481$'biorep' <- "pr.481"
counts$pr.484$'biorep' <- "pr.484"
```
## Create list of freemuxlet raw output data
```{r}
freemux_dir <- paste0(data_dir, "freemuxlet/raw_data/")
names <- 
  grep(
    dir(
      path = freemux_dir             #list all files, unfortunately includes .ini files
      ),
    pattern = "(kc|pr)(_\\d+)+", #pattern to search for, excludes .ini files
    value = T                    #return matches
    )
files <-
  paste0(
    freemux_dir, 
    names
  )
# Read in freemuxlet dataframes, specifying the datatype in each column
freemux.clusters <- lapply(files, readr::read_tsv, col_types = "iciiffdfddddfdfddfdd")
names(freemux.clusters) <- names

# Optional? 'QC' freemux output?
```
## Custom function to add freemuxlet cluster assignment to Seurat MetaData
```{r}
# Add freemuxlet cluster assignment to Seurat Metadata
AddFreemuxletMetaData <- function(seu, freemux.data, col.name) {
  seu.barcodes <- tibble (
    BARCODE = rownames(seu@meta.data)) # get barcodes list from Seurat object
  join <- left_join(                      # join freemuxlet assignment with barcodes
    seu.barcodes,
    freemux.data %>%
      dplyr::select(BARCODE, BEST.GUESS) %>%
      mutate(
        BEST.GUESS = fct_recode(          # Refactor freemuxlet assignments
          BEST.GUESS,
          "1" = "1,1",
          "0" = "0,0",
          "doublet" = "1,0"
        )
      )
  )
  seu <- AddMetaData(seu,                 # Add MetaData to Seurat object
                     join %>% pull(BEST.GUESS),
                     col.name = col.name)
  return(seu)                             # Return augmented Seurat object
}
```
## Add freemuxlet cluster assignment to Seurat objects
Error output for pr.478 because it is the only dataset with no assigned doublets. Output appears normal otherwise
```{r}
mapped <-
  map2(counts,                              # List of Seurat objects
       freemux.clusters,                    # List of freemuxlet raw data
       AddFreemuxletMetaData,               # Function to call
       col.name = "freemuxlet.assignments") # Every metadata column has same name

rm(counts)       # Remove data
rm(freemux.clusters) # Remove data
```
## Merge Seurat objects and remove doublets
```{r}
# Merge, adding cell id information because barcodes can and have been duplicated across experiments
kc.seu <- merge(x = mapped[[1]],                # Seurat object
             y = mapped[(2:4)],             # List
             add.cell.ids = names(mapped)[1:4])   # Add cell IDs using names
# Remove doublets
kc.seu <- subset(x = kc.seu,
              subset = freemuxlet.assignments == "doublet",
              invert = T)

# Merge, adding cell id information becaue barcodes can and have been duplicated across experiments
pr.seu <- merge(x = mapped[[5]],                # Seurat object
             y = mapped[(6:7)],             # List
             add.cell.ids = names(mapped)[5:7])   # Add cell IDs using names
pr.seu <- subset(x = pr.seu,
              subset = freemuxlet.assignments == "doublet",
              invert = T)
```

# KC QC
```{r}
# Convert to SingleCellExperiment for QC
sce <- as.SingleCellExperiment(kc.seu)
# Factorize the batch variable
sce$batch <- as.factor(sce$batch)
```

## Droplet Calling
Skipping droplet calling since 10x already filters that under Cell Ranger 4.0 based on Aaron Lun's EmptyDroplets().

## Calculate QC Metrics
Increased NMADs to 4 because NMAD of 3.0 looks too restrictive for KC data. Looks like an appropriate compromise based on QC plots.
```{r}
# Find mitochondrial genes
is.mito <- grepl("^MT-", rownames(sce))
# Add per cell QC metrics to the object
sce <- addPerCellQC(sce, subsets = list(Mito=is.mito))

# Create QC data frame
qc.df <- perCellQCMetrics(sce, subsets = list(Mito=is.mito))
# Reasons for exclusion, considering batch
batch.reasons <- quickPerCellQC(qc.df, batch=sce$batch,
                                sub.fields = "subsets_Mito_percent",
                                nmads = 4)
colSums(as.matrix(batch.reasons))

# Add discard metadata to object
sce$discard <- batch.reasons$discard
discard <- batch.reasons$discard
```

QC plots.
```{r}
plotColData(sce, x="sum", y="subsets_Mito_percent", colour_by = "discard") + scale_x_log10()
plotColData(sce, x="detected", y="subsets_Mito_percent", colour_by = "discard") + scale_x_log10()
```

```{r}
plotColData(sce, x="ident", y="sum", colour_by="discard") + 
        scale_y_log10() + ggtitle("Total count")
plotColData(sce, x="ident", y="detected", colour_by="discard") + 
        scale_y_log10() + ggtitle("Detected features")
plotColData(sce, x="ident", y="subsets_Mito_percent", 
        colour_by="discard") + ggtitle("Mito percent")
```
Exploratory investigation of variance explained by metadata.
```{r}
#vars <- getVarianceExplained(logNormCounts(sce), 
#    variables=c("ident", "subsets_Mito_percent", "sum", "detected"))
#plotExplanatoryVariables(vars)
```
Subset to QC-filtered cells.
```{r}
# Keeping the columns we DON'T want to discard.
filtered <- sce[,!discard]
dropped <- as.numeric(summary(discard)["TRUE"])
kept <- as.numeric(summary(discard)["FALSE"])
print(paste0(dropped, " cells were filtered, yielding ", kept, " remaining cells"))
```
Identify if a rare cell type might have been discarded.
```{r}
lost <- calculateAverage(counts(sce)[,!discard])
kept <- calculateAverage(counts(sce)[,discard])

logged <- cpm(cbind(lost, kept), log=TRUE, prior.count=2)
logFC <- logged[,1] - logged[,2]
abundance <- rowMeans(logged)
```

Blue points are mitochondrial genes and are expected to be downregulated. Exclusion of a cell type may be indicated by overexpression on this plot, of which there does not appear to be strong overexpression.
```{r}
plot(abundance, logFC, xlab="Average count", ylab="Log-FC (lost/kept)", pch=16)
points(abundance[is.mito], logFC[is.mito], col="dodgerblue", pch=16)
```
Even at NMAD = 4, CSH1 is differentially underexpressed, indicates omission of syncytiotrophoblasts, possibly correct since syncytiotrophoblast may have been differentially technically less sound.
```{r}
# Pull out differentially regulated genes
logFC[logFC < -2]

plotExpression(sce, features = "CSH1", x = "subsets_Mito_percent")
plotExpression(filtered, features = "CSH1", x = "subsets_Mito_percent")

plotExpression(sce, features = "CSH1", x = "total")
plotExpression(filtered, features = "CSH1", x = "total")

plotExpression(sce, features = "CSH1", x = "detected")
plotExpression(filtered, features = "CSH1", x = "detected")
```

Convert to Seurat object for further preprocessing and analysis and save.
```{r}
kc <- as.Seurat(filtered)
#saveRDS(kc, paste0(data_dir, "placenta_scRNA_seq_kc_filtered_", Sys.Date(), ".rda"))
```

Explore batch effect in KC samples. There doesn't appear to be much of a batch effect.
```{r}
# Should normalize and find variable features in each dataset separately
kc <- NormalizeData(object = kc, normalization.method = "LogNormalize", scale.factor = 10000)
kc <- FindVariableFeatures(object = kc)
kc <- ScaleData(kc)
kc <- RunPCA(kc, features = VariableFeatures(object = kc))
ElbowPlot(object = kc, ndims = 100)
DimPlot(kc, reduction = "pca", group.by = 'orig.ident')
DimPlot(kc, reduction = "pca", group.by = 'biorep')
```

```{r}
# Rename Samples KC_40_1 and KC_40_2 have been renamed 1A and 1B; KC_42_1 and KC_42_2 <- 2A and 2B.
#levels(All@active.ident) <- c("Placenta 1A", "Placenta 1B", "Placenta 2A", "Placenta 2B")
```

# Pique-Regi data QC

```{r}
# Convert to SingleCellExperiment for QC
sce <- as.SingleCellExperiment(pr.seu)
# Factorize the batch variable
sce$batch <- as.factor(sce$batch)
# Clean-up memory
# rm(seu)
```

## Droplet Calling
Skipping droplet calling since 10x already filters that under Cell Ranger 4.0 based on Aaron Lun's EmptyDroplets().

## Calculate QC Metrics
Increased NMADs to 3.5 because NMAD of 3.0 looks too restrictive for KC data and 4.0 looked to permissive for PR data. Looks like an appropriate compromise based on QC plots.
```{r}
# Find mitochondrial genes
is.mito <- grepl("^MT-", rownames(sce))
# Add per cell QC metrics to the object
sce <- addPerCellQC(sce, subsets = list(Mito=is.mito))

# Create QC data frame
qc.df <- perCellQCMetrics(sce, subsets = list(Mito=is.mito))
# Reasons for exclusion, considering batch
batch.reasons <- quickPerCellQC(qc.df,
                                sub.fields = "subsets_Mito_percent",
                                nmads = 3) # default nmads is 3
colSums(as.matrix(batch.reasons))

# Add discard metadata to object
sce$discard <- batch.reasons$discard
discard <- batch.reasons$discard
```

QC plots.
```{r}
plotColData(sce, x="sum", y="subsets_Mito_percent", colour_by = "discard") + scale_x_log10()
plotColData(sce, x="detected", y="subsets_Mito_percent", colour_by = "discard") + scale_x_log10()
```

```{r}
plotColData(sce, x="ident", y="sum", colour_by="discard") + 
        scale_y_log10() + ggtitle("Total count")
plotColData(sce, x="ident", y="detected", colour_by="discard") + 
        scale_y_log10() + ggtitle("Detected features")
plotColData(sce, x="ident", y="subsets_Mito_percent", 
        colour_by="discard") + ggtitle("Mito percent")
```
Exploratory investigation of variance explained by metadata.
```{r}
#vars <- getVarianceExplained(logNormCounts(sce), 
#    variables=c("ident", "subsets_Mito_percent", "sum", "detected"))
#plotExplanatoryVariables(vars)
```
Subset to QC-filtered cells.
```{r}
# Keeping the columns we DON'T want to discard.
filtered <- sce[,!discard]
dropped <- as.numeric(summary(discard)["TRUE"])
kept <- as.numeric(summary(discard)["FALSE"])
print(paste0(dropped, " cells were filtered, yielding ", kept, " remaining cells"))
```
Identify if a rare cell type might have been discarded.
```{r}
lost <- calculateAverage(counts(sce)[,!discard])
kept <- calculateAverage(counts(sce)[,discard])

logged <- cpm(cbind(lost, kept), log=TRUE, prior.count=2)
logFC <- logged[,1] - logged[,2]
abundance <- rowMeans(logged)
```

Blue points are mitochondrial genes and are expected to be downregulated. Exclusion of a cell type may be indicated by overexpression on this plot, of which there does not appear to be strong overexpression.
```{r}
plot(abundance, logFC, xlab="Average count", ylab="Log-FC (lost/kept)", pch=16)
points(abundance[is.mito], logFC[is.mito], col="dodgerblue", pch=16)
```
No clear cell types omitted based on my prior knowledge.
```{r}
logFC[logFC < -2]
```

Convert to Seurat object for further preprocessing and analysis and save.
```{r}
pr <- as.Seurat(filtered)
```

Explore batch effect in PR samples. There may be some batch effect.
```{r}
# Should normalize and find variable features in each dataset separately
pr <- NormalizeData(object = pr, normalization.method = "LogNormalize", scale.factor = 10000)
pr <- FindVariableFeatures(object = pr)
pr <- ScaleData(pr)
pr <- RunPCA(pr, features = VariableFeatures(object = pr))
ElbowPlot(object = pr, ndims = 100)
DimPlot(pr, reduction = "pca", group.by = 'orig.ident')
```

# TODO
Clean-up merged object, looks like merging causes duplicate metadata columns
```{r}
merged <- merge(kc, pr)
#saveRDS(merged, paste0(data_dir, "kc_pr_qc_merged_", Sys.Date(), ".rda"))
```