---
title: "GSE75010 Cell Type Proportion Prep"
author: "Kyle Campbell"
date: "1/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(factoextra)
library(tidyverse)
library(betareg)
library(boot)
library(psych)
library(readxl)

# BiocPackages to use with GEOquery
library(Biobase)
library(BiocGenerics)
library(GEOquery)

data_dir <- paste0(here("data"), "/")
results_dir <- paste0(here("results", "GSE75010"), "/")
```
Leavy & Cox et al. previously analyzed the 7 microarray studies in:
Large Scale Aggregate Microarray Analysis Reveals Three Distinct Molecular Subclasses of Human Preeclampsia

## GEO Query
Pull GEO accession as GSE Matrix file
```{r}
#gse <- getGEO("GSE75010", GSEMatrix = TRUE)
#saveRDS(gse, paste0(data_dir, "GSE75010_GSEMatrix.rda"))
gse <- readRDS(paste0(data_dir, "GSE75010_GSEMatrix.rda"))
head(gse[[1]])
# Detailed phenotype data
#head(gse$GSE75010_series_matrix.txt.gz@phenoData@data)
```

Pull relevant information from the Cox PE dataset. Overlapping covariates between the Cox datasets are fetal sex and gestational age. GSE75010 additionally contains ethnicity, gestational age in days, attempted vaginal delivery, and mode of delivery whereas the aggregated microarrays contain occurrence of labor alone and nationality.
```{r}
# Pull SAMPLE_XXXXX,  PE vs. non-PE, GA, and sex from Cox dataset
cox <- gse$GSE75010_series_matrix.txt.gz@phenoData@data[,c("description", "characteristics_ch1", "characteristics_ch1.18", "characteristics_ch1.20")]
cox <- rownames_to_column(cox, "gsm")
cox$batch <- "GSE75010"
colnames(cox) <- c("gsm", "sample", "phenotype", "ga", "sex", "batch")
cox <- cox %>%
  dplyr::select(sample, phenotype, batch, ga, sex)
head(cox)
```

Recode Cox dataset to match microarray data.
```{r}
# Recode diagnosis: non-PE to control
cox$phenotype <- sapply(cox$phenotype,
       function(x) gsub(x, pattern = ".*[\\-]PE",
       replacement = "C"
       ))
# Recode diagnosis: PE to PE
cox$phenotype <- sapply(cox$phenotype,
       function(x) gsub(x, pattern = ".*[^\\-]PE",
       replacement = "PE"
       ))
# Recode ga (week): xx to xx weeks and convert to numeric
cox$ga <- sapply(cox$ga,
       function(x) gsub(x, pattern = ".*[^0-9]{2}",
       replacement = ""
       ))
cox$ga <- as.numeric(cox$ga)
# Recode sex
cox$sex <- sapply(cox$sex,
       function(x) gsub(x, pattern = ".*[^F|M]",
       replacement = ""
       ))
head(cox)
```

Pull relevant information from the microarray aggregation done by Cox group.
```{r}
array <- read_excel(paste0(data_dir, "journal.pone.0116508.s014.xlsx"), col_names = TRUE)
head(array)
array <- array %>%
  dplyr::select(Sample, Phenotype, Batch, GA, `Fetal Sex`)
colnames(array) <- c("sample", "phenotype", "batch", "ga", "sex")
head(array)
```

Merging the two metadata tables
```{r}
concat <- rbind(cox, array)
head(concat)
dim(concat)
```

Import CIBERSORT cell type fraction imputation data.
```{r}
frac <- read_excel(paste0(data_dir, "cibersortx/output/CIBERSORTx_Job36_Results_no_min_expr_2000_max_genes.xlsx"))
colnames(frac)[1] <- "sample"
head(frac)
dim(frac)
```

Merge the two phenotype data with the cell type fraction data and factorize appropriate variables. Note typo in ST cell type.
```{r}
merged <- left_join(concat, frac)
merged$phenotype <- as.factor(merged$phenotype)
merged$batch <- as.factor(merged$batch)
merged$sex <- as.factor(merged$sex)

#colnames(merged) <- tolower(colnames(merged))
#colnames(merged) <- str_replace_all(colnames(merged),
#                                "\\+",
#                                "_")
#colnames(merged) <- str_replace_all(colnames(merged),
#                                "\\s",
#                                "_")
#colnames(merged) <- str_replace_all(colnames(merged),
#                                "\\-",
#                                "_")

# Fix previous typos
colnames(merged)[colnames(merged)=="Fetal CD14+ Monoctyes"] <- "Fetal CD14+ Monocytes"
colnames(merged)[colnames(merged)=="Maternal FCGR3A+ Monoctyes"] <- "Maternal FCGR3A+ Monocytes"
colnames(merged)[colnames(merged)=="Maternal CD14+ Monotyes"] <- "Maternal CD14+ Monocytes"

head(merged)
dim(merged)
#saveRDS(merged, paste0(data_dir, "GSE75010_with_abundances", Sys.Date(), ".rda"))
#saveRDS(merged, paste0(data_dir, "GSE75010_with_abundances2021-01-01.rda"))
#merged <- readRDS(paste0(data_dir, "GSE75010_merged_abundances2020-10-16.rda"))
```

Summary stats
```{r}
describe(merged$`P-value`)
describe(merged$Correlation)
describe(merged$RMSE)
```

```{r}
describe(merged$`Fetal Cytotrophoblasts`)
describe(merged$`Fetal Syncytiotrophoblast`)
describe(merged$`Fetal Naive CD8+ T Cells`)
```

```{r}
describe(merged$`Maternal Naive CD8+ T Cells`)
describe(merged$`Maternal Natural Killer Cells`)
describe(merged$`Maternal Plasma Cells`)
```

