---
title: "Freemuxlet results"
author: "Kyle Campbell"
date: "02/13/2022"
output: html_document
---

```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(ggridges)

knitr::opts_chunk$set(echo = TRUE)

data_dir <- paste0(here("data", "freemuxlet", "raw_data"), "/")
```

Get names
```{r}
tenx_out <- paste0(here("data", "tenx_outs"), "/")
# Pull full file paths
names <- 
  grep(
    dir(
      path = tenx_out             #list all files, unfortunately includes .ini files
      ),
    pattern = "(kc|pr|tsang)", #pattern to search for, excludes .ini files
    value = T                    #return matches
    )
```

```{r}
# Pull full file paths
files <- paste0(data_dir, dir(path = data_dir))
# Subtracting the desktop.ini file
files <- files[-1]
# Read in freemuxlet dataframes, specifying the datatype in each column
clusters <- lapply(files, readr::read_tsv, col_types = "iciiffdfddddfdfddfdd")
# Rename the list of dataframes with sample ID
names(clusters) <- names
```

```{r}
# Function to summarize guesses
summarize_freemuxlet_data <- function(freemux.data) {
  # Summarize droplet assignment results
  summary(as.factor(freemux.data$BEST.GUESS))
}

# Create nested tibble
df <- tibble(
  sample = names,
  data = clusters
)

# Unnest the data to add sample labels
all <- df %>% unnest(data)

# Summarize best guess results
all %>%
  group_by(sample, BEST.GUESS) %>%
  tally() %>%
  View

#saveRDS(all, paste0(data_dir, "freemuxlet_assignments_", Sys.Date(), ".rda"))
#write_csv(all, paste0(data_dir, "freemuxlet_assignments_", Sys.Date(), ".csv"))
```

```{r}
test <- 
  all %>%
  mutate(
        BEST.GUESS = fct_recode(          # Refactor freemuxlet assignments
          BEST.GUESS,
          "1" = "1,1",
          "0" = "0,0",
          "doublet" = "1,0"
        ))
```


Cluster represents freemuxlet clusters assigned from entire 1-22, X, Y 1000Genomes .vcf reference.
Cluster.minaf represents 1-22, X 1000Genomes genotype data only, filtering rare variants with MAF < 10%

```{r}
cluster <- read_tsv(here("data", "freemuxlet", "pr478clusters.txt"))
cluster.minaf <- read_tsv(here("data", "freemuxlet", "pr478clusters.minAF.txt"))

# Sort both by BARCODE
cluster <- 
  cluster %>%
  arrange(BARCODE)
cluster.minaf <-
  cluster.minaf %>%
  arrange(BARCODE)

# Are there differences in best guess?
summary(cluster$BEST.GUESS == cluster.minaf$BEST.GUESS)

all.best <- as.factor(cluster$BEST.GUESS)
summary(all.best)
minaf.best <- as.factor(cluster.minaf$BEST.GUESS)
summary(minaf.best)
```
Create new variables to investigate differences in assignment; no issues
```{r}
all <- cluster
min <- cluster.minaf

mismatch <- !(cluster$BEST.GUESS == cluster.minaf$BEST.GUESS)

all$mismatch <- mismatch
min$mismatch <- mismatch

all$filtered <- FALSE
min$filtered <- TRUE

data <- rbind(all, min) %>%
  arrange(BARCODE)
data.mismatch <- data %>%
  filter(mismatch == T)
```

```{r}
data <- data.frame(
  all <- cluster$DIFF.LLK.BEST.NEXT,
  min <- cluster.minaf$DIFF.LLK.BEST.NEXT
)
```
