---
title: "single-cell-analysis"
author: "Kyle Campbell"
date: "1/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

#install.packages("package_name",  repos="http://cran.us.r-project.org") # Only this mirror worked for multiple packages
library(devtools)
library(here)
library(ggrepel)
library(ggpubr)
library(gprofiler2)
library(knitr)
library(openxlsx)
library(tidyverse)
library(Seurat)
#remotes::install_github('satijalab/seurat-wrappers')
library(SeuratWrappers)
library(cowplot)
library(rgl)
library(EnvStats)
library(reticulate)
library(clustree)
library(Matrix)

# Bioconductor packages
#remotes::install_github('chris-mcginnis-ucsf/DoubletFinder') #Install DoubletFinder and depends
library(batchelor)
library(edgeR)
library(scater)
# Set seed since this pipeline includes stochastic algorithms
set.seed(1)
```

Load cleaned Seurat data
```{r}
#seu <- readRDS(here("data", "analytic", "2022-04-18_all_merged_pruned_batch_mnn.rda"))
# Analytic Tsang resubmit
seu <- readRDS(here("data", "analytic", "2022-05-01_all_merged_pruned_batch_mnn.rda"))
seu$cell.type <- seu$cell.type %>% factor()
Idents(seu) <- seu$cell.type
```

QC metrics by replicate, TODO: relabel orig.ident for nicer graphs
```{r}
VlnPlot(seu, features = "subsets_Mito_percent", pt.size = .1, group.by = "orig.ident") + NoLegend() + 
  theme(axis.text.x = element_text(size = 8)
        ) +
  ggtitle("Percentage of Reads Mapping to Mitochondrial Genes by Replicate")
#ggsave(paste0(results_dir, "qc_replicate_mito_", Sys.Date(), ".png"))

VlnPlot(seu, features = "nCount_RNA", pt.size = .1, group.by = "orig.ident") + NoLegend() + 
  theme(axis.text.x = element_text(size = 8)
        ) +
  ggtitle("Total RNA Molecules Per Cell by Replicate")
#ggsave(paste0(results_dir, "qc_replicate_total_RNA_", Sys.Date(), ".png"))

VlnPlot(seu, features = "nFeature_RNA", pt.size = .1, group.by = "orig.ident") + NoLegend() + 
  theme(axis.text.x = element_text(size = 8)
        ) +
  ggtitle("Detected Genes Per Cell by Replicate")
#ggsave(paste0(results_dir, "qc_replicate_total_genes_", Sys.Date(), ".png"))
```

### Supplementary Figure - Marker gene expression by cell type cluster

REORDER factor levels for Idents (it's this way on some of the supplementary plots)
```{r}
# Indicator for whether non-immune fetal cell
seu$immune <- ifelse(seu$cell.type %in% c("Fetal CD14+ Monocytes", "Fetal CD8+ Activated T Cells", "Fetal Naive CD4+ T Cells", "Fetal Naive CD8+ T Cells", "Fetal Natural Killer T Cells", 
"Fetal B Cells", "Fetal GZMK+ Natural Killer", "Fetal Memory CD4+ T Cells", "Fetal Plasmacytoid Dendritic Cells", 
"Fetal GZMB+ Natural Killer", "Maternal CD8+ Activated T Cells", 
"Maternal Naive CD4+ T Cells", "Maternal FCGR3A+ Monocytes", 
"Maternal CD14+ Monocytes", "Maternal Natural Killer Cells", 
"Maternal B Cells", "Maternal Plasma Cells", "Maternal Naive CD8+ T Cells"), T, F)

seu.immune <- subset(seu, immune == T)
seu.nonimmune <- subset(seu, immune == F)


DotPlot(seu.nonimmune, features = c("KRT7", "STMN1", "HLA-G", "PSG4","CD163", "PECAM1", "COL1A1", "TAGLN", "LUM", "HBA"), scale.by = "size") + RotatedAxis() 

genes.seurat <- rownames(seu@assays$RNA@data) %>% as.data.frame()
```

```{r}
base.size = 24
#dotplot.nonimmune <- DotPlot(seu.nonimmune, features = c("KRT7", "STMN1", "HLA-G", "PSG4","CD163", "PECAM1", "COL1A1", "TAGLN", "LUM", "HBB"), scale.by = "size") + theme_bw(base_size = base.size) + RotatedAxis() 
#dotplot.immune <- DotPlot(seu.immune, features = c("CD79A", "IGHA1", "XBP1", "NKG7", "CD3E", "CD4", "IL2", "CCR7", "CD8A", "GZMB", "GZMK", "FCGR3A", "CD14", "FLT3", "ITM2C"), scale.by = "size") + theme_bw(base_size = base.size, ) + RotatedAxis()
#dotplot.legend <- get_legend(dotplot.immune)

dotplot.nonimmune <- DotPlot(seu.nonimmune, features = c("KRT7", "STMN1", "HLA-G", "PSG4","CD163", "PECAM1", "COL1A1", "TAGLN", "LUM", "HBB"), scale.by = "size") + theme_bw(base_size = base.size) + RotatedAxis() 
dotplot.immune <- DotPlot(seu.immune, features = c("CD79A", "IGHA1", "XBP1", "NKG7", "CD3E", "CD4", "IL2", "CCR7", "CD8A", "GZMB", "GZMK", "FCGR3A", "CD14", "FLT3", "ITM2C"), scale.by = "size") + theme_bw(base_size = base.size) + theme(legend.box = "horizontal") + RotatedAxis()
dotplot.legend <- get_legend(dotplot.immune)

panel <- ggarrange(dotplot.nonimmune, dotplot.immune, ncol = 2, nrow = 1, labels = "AUTO", font.label = list(size = 30, face = "bold", color = "black"), legend = "bottom", legend.grob = dotplot.legend)
#ggexport(panel, filename = here("results", "seurat", paste0(Sys.Date(), "_dotplot_marker_genes.png")), width = 1980, height = 1080)
```


plasma cell vs B cell
Huge overepxression of immunoglobulins compared to B cells
```{r}
plasma.cell <- FindMarkers(seu, ident.1 = "Maternal Plasma Cells", ident.2 = "Maternal B Cells")
```


```{r}
DimPlot(seu, label = T, repel = T) + NoLegend()
FeaturePlot(object = seu,
            features = c("HLA-DRA"),
            cols = c("lightgrey", "blue"),
            pt.size = .1)
FeaturePlot(object = seu,
            features = c("CLEC4C", "IL3RA"),
            cols = c("lightgrey", "blue"),
            pt.size = .1)
FeaturePlot(object = seu,
            features = c("GZMB", "SERPINF1"),
            cols = c("lightgrey", "blue"),
            pt.size = .1)
FeaturePlot(object = seu,
            features = c("IL3RA"),
            cols = c("lightgrey", "blue"),
            pt.size = .1)
FeaturePlot(object = seu,
            features = c("SMIM24"),
            cols = c("lightgrey", "blue"),
            pt.size = .1)
```

```{r}
x.vlnplot <- function(gene_vector) {
  map(gene_vector, ~ VlnPlot(seu.immune, .x) + NoLegend())
}
```

```{r}
x.vlnplot(c("NKG7", "CD3E", "CD4", "S100A4", "CCR7", "IL2RA", "FOXP3", "CD8A"))
```


```{r}
x.dotplot <- function(gene_vector) {
  DotPlot(seu.immune, features = gene_vector) + RotatedAxis()
}
```
https://doi.org/10.3389/fimmu.2019.01222 also indicates IL-7R+ and FLT3+ precursor population can give rise to B cells or pDCs, consistent with some pDC genes being expressed in both
```{r}
x.dotplot(c("SERPINF1", "AK128525", "IGJ", "ITM2C", "PLD4", "CCDC50"))
x.dot
```


```{r}
x.vlnplot(c("GZMB", "SELPLG", "DERL3", "PTPRCAP", "BCL11A", "LAMP5", "SLA2", "NRP1", "SIDT1", "TCF4", "SLC15A4", "IRF7", "TRAF4"))
```


### Supplementary Figure - QC metrics by cell type cluster 

QC metrics by cluster
```{r}
base.size <- 24
x.lab.size <- 20
x.lab.hjust <- 0.75

cluster.rna <- VlnPlot(seu, features = "nCount_RNA", pt.size = .1) + theme_bw(base_size = base.size) + NoLegend() + coord_flip() + 
  theme(axis.text.y = element_blank(), axis.title.y = element_blank(), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(size = x.lab.size, hjust = x.lab.hjust)) + 
  ggtitle("Total unique RNA molecules") 

cluster.rna.legend <- VlnPlot(seu, features = "nCount_RNA", pt.size = .1) + theme_bw(base_size = base.size) + coord_flip() + 
  theme(legend.text = element_text(size = 32, hjust = 1), legend.spacing = unit(1, "npc")) + 
  ggtitle("Total unique RNA molecules") +  guides(fill = guide_legend(ncol = 1, keyheight = 2.5, reverse = T))
legend <- get_legend(cluster.rna.legend)
#as_ggplot(legend)

cluster.gene <- VlnPlot(seu, features = "nFeature_RNA", pt.size = .1) + theme_bw(base_size = base.size) + NoLegend() + 
  theme(axis.text.y = element_blank(),  axis.title.y = element_blank(), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(size = x.lab.size, hjust = x.lab.hjust)) +
  ggtitle("Detected genes") + coord_flip()

cluster.mito <- VlnPlot(seu, features = "subsets_Mito_percent", pt.size = .1) + theme_bw(base_size = base.size) + NoLegend() + 
  theme(axis.text.y = element_blank(), axis.title.y = element_blank(), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(size = x.lab.size, hjust = x.lab.hjust)) +
  ggtitle("% Mitochondrial genes") + coord_flip()
```

```{r}
panel <- ggarrange(cluster.rna, cluster.gene, cluster.mito, ncol = 3, nrow = 1, labels = "AUTO", font.label = list(size = 30, face = "bold", color = "black"), legend = "left", legend.grob = legend)
#ggexport(panel, filename = here("results", "seurat", paste0(Sys.Date(), "_cluster_qc_vlns.png")), width = 1980, height = 1080)
```

Seurat's built-in version that I didn't use.
```{r, eval = F}
seurat.cluster.qc.plot <- VlnPlot(
  seu,
  features = c("nCount_RNA", "nFeature_RNA", "subsets_Mito_percent"),
  cols = NULL,
  pt.size = NULL,
  idents = NULL,
  sort = FALSE,
  assay = NULL,
  group.by = NULL,
  split.by = NULL,
  adjust = 1,
  y.max = NULL,
  same.y.lims = FALSE,
  log = FALSE,
  ncol = NULL,
  slot = "data",
  split.plot = FALSE,
  stack = FALSE,
  combine = FALSE,
  fill.by = "ident",
  flip = FALSE
) + NoLegend() + theme(axis.title.y = element_blank(), axis.title.x = element_blank())
#ggsave(plot = seurat.cluster.qc.plot, filename = here("results", "seurat", paste0(Sys.Date(), "_seurat_cluster_qc_vlns.png")))
```

### Supplementary Figure - XIST expression by biological replicate
Fetal sex check using XIST as a biomarker for female sex
```{r}
seu <- NormalizeData(seu)
seu$fetal <- seu$fetal %>% as.factor()
fetal <- subset(seu, subset = fetal == "Fetal")

biorep.expr <- AverageExpression(fetal, group.by = c("fetal", "biorep"))
biorep.expr <- biorep.expr$RNA %>% t() %>% as.data.frame()
biorep.expr <- rownames_to_column(biorep.expr, var = "biorep")
#View(head(biorep.expr))

biorep.expr$biorep <- factor(biorep.expr$biorep)
levels(biorep.expr$biorep)
levels(biorep.expr$biorep) <- c("Sample 1", "Sample 2", "Sample 3", "Sample 4", "Sample 5", "Sample 6", "Sample 7", "Sample 8", "Sample 9")

ggplot(data = biorep.expr, aes(x = biorep, y = XIST)) + geom_point() + labs(x = "Biological Replicate") + ggtitle("Average XIST expression in fetal cells by biological replicate") + theme_minimal()

#ggsave(here("results", "seurat", paste0(Sys.Date(), "xist_sex_expression.png")))
```

```{r}
DimPlot(seu, label = T, label.size = label.size, repel = T) + NoLegend() + theme(axis.text = element_text(size = axis.text.size), axis.title = element_text(size = axis.title.size))
```

### Figure 1
UMAP plots for Figure 1
```{r}
# Set label size for Seurat DimPlots
label.size <- 9
axis.text.size <- 24
axis.title.size <- 28

all.umap <- DimPlot(seu, label = T, label.size = label.size, repel = T) + NoLegend() + theme(axis.text = element_text(size = axis.text.size), axis.title = element_text(size = axis.title.size))
fetal.umap <- subset(seu, subset = fetal == "Fetal") %>% DimPlot(label = T, label.size = label.size, repel = T) + NoLegend() + theme(axis.text = element_text(size = axis.text.size), axis.title = element_text(size = axis.title.size))
maternal.umap <- subset(seu, subset = fetal == "Maternal") %>% DimPlot(label = T, label.size = label.size, repel = T) + NoLegend() + theme(axis.text = element_text(size = axis.text.size), axis.title = element_text(size = axis.title.size))

umap.panel <- ggarrange(
  all.umap,
  ggarrange(
    fetal.umap,
    maternal.umap,
    ncol = 2,
    labels = c("B", "C"),
    font.label = list(size = 36, face = "bold", color = "black")
  ),
  nrow = 2,
  labels = "A",
  font.label = list(size = 36, face = "bold", color = "black")
)

ggexport(umap.panel, filename = here("results", "seurat", "umap", paste0(Sys.Date(), "_umap_multipanel.png")), width = 1920, height = 1080)

DimPlot(seu, label = T, repel = T) + NoLegend()
#ggsave(filename = here("results", "seurat", paste0(Sys.Date(), "_all_umap.png")))
```

VEGFR Receptor expression by cluster
```{r}
VlnPlot(seu, features = "KDR") + NoLegend()
VlnPlot(seu, features = "FLT1") + NoLegend()
```

Looking at Preeclampsia-related genes from "Hypoxia in the pathogenesis of preeclampsia" by Keiichi Matsubara
```{r}
VlnPlot(seu, features = "ENG") + NoLegend()
VlnPlot(seu, features = "LEP") + NoLegend()

VlnPlot(seu, features = "HIF1A") + NoLegend()
#VlnPlot(seu, features = "HIF1B") + NoLegend()
VlnPlot(seu, features = "TGFB3") + NoLegend()

FeaturePlot(seu, features = "HIF1A") + NoLegend()

hif <- grepl("HIF", rownames(seu))
hifs <- rownames(seu)[hif]
  
VlnPlot(seu, features = "CD9") + NoLegend()
#ggsave(paste0(results_dir, "FLT1_expr_by_cluster_", Sys.Date(), ".png"))
```

Percentage of fetal cells
```{r}
mf <- seu$fetal %>% as.factor() %>% summary()
percent.fetal <- mf[1]/(mf[1]+mf[2])
```

Paint by fetal/maternal status
```{r}
legend.text.size <- 24
umap.pt.size <- 1.1
umap.mf <- DimPlot(seu, group.by = 'fetal', pt.size = umap.pt.size) + ggtitle("") + theme(legend.text = element_text(size = legend.text.size))
DimPlot(seu, group.by = 'fetal') + ggtitle(paste0(round(percent.fetal, 4)*100, "% fetal cells"))
ggsave(here(paste0("seu_group_by_fetal_status_", Sys.Date(), ".png")))
```

```{r}
#seu$batch <- factor(seu$batch, labels = c("KC", "PR"))
```

Paint by batch
```{r}
DimPlot(seu, group.by = 'batch') + ggtitle("By batch")
#ggsave(here("results", paste0("seu_group_by_batch_", Sys.Date(), ".png")))
```

```{r}
DimPlot(seu, group.by = 'batch')
#ggsave(paste0(results_dir, "seu_group_by_batch_legend_", Sys.Date(), ".png"))
```

```{r}
seu.graph <- seu
seu.graph$biorep <- factor(seu.graph$biorep, labels = c("Sample 1" , "Sample 2" , "Sample 3", "Sample 4", "Sample 5", "Sample 6", "Sample 7", "Sample 8", "Sample 9"))

umap.biorep <- DimPlot(seu.graph, group.by = 'biorep', pt.size = umap.pt.size) + ggtitle("") + theme(legend.text = element_text(size = legend.text.size))

DimPlot(seu.graph, group.by = 'biorep') + ggtitle("")
#ggsave(here("results", "seurat", paste0(Sys.Date(), "seu_group_by_biorep.png")))
rm(seu.graph)
```

Subset to and paint by technical replicates
```{r}
kc.40 <- subset(seu, subset = biorep == "kc.40")
Idents(kc.40) %>% levels
kc.40$orig.ident <- factor(kc.40$orig.ident, labels = c("Sample 1A", "Sample 1B"))


umap.sample1 <- DimPlot(kc.40, group.by = "orig.ident", pt.size = umap.pt.size) + ggtitle("") + theme(legend.text = element_text(size = legend.text.size))

DimPlot(kc.40, group.by = "orig.ident")
#ggsave(paste0(results_dir, "seu_group_by_kc40_technical_legend_", Sys.Date(), ".png"))
rm(kc.40)
```

Average expression by cell type cluster and run corrtest.
```{r}
# Subset to the technical replicate
kc.40.1 <- subset(seu, subset = orig.ident == "kc.40.1") %>% NormalizeData()
# Average expression by cell type
expr.1a <- AverageExpression(kc.40.1, group.by = "cell.type")
# Clean-up the returned object
a1 <- expr.1a$RNA %>% as.data.frame()

kc.40.2 <- subset(seu, subset = orig.ident == "kc.40.2")  %>% NormalizeData()
#cluster.expr.40.2 <- flatten(AverageExpression(kc.40.2)) %>% as.numeric
expr.1b <- AverageExpression(kc.40.2, group.by = "cell.type")
b1 <- expr.1b$RNA %>% as.data.frame()

# Subset a1 and b1 to the cell types that are contained in both
common <- intersect(a1 %>% colnames, b1 %>% colnames)

a1 <- a1 %>% dplyr::select(all_of(common))
b1 <- b1 %>% dplyr::select(all_of(common))
```


```{r}
corr.test.res <- map2(a1,
                      b1,
                      ~ cor.test(.x, .y, method = "pearson"))

corr.test.pvals <- map(corr.test.res,
                   ~ .x['p.value'])

pvals <- flatten(corr.test.pvals) %>% flatten

corr.res <- map2(a1,
     b1,
     ~ cor(.x, .y))

cor.df <- corr.res %>% data.frame %>% t
colnames(cor.df) <- "correlation"
cor.df <- as.data.frame(cor.df)

cor.df$cell.type <- rownames(cor.df)
```

```{r}
sample1.avg.cor <- cor.df$correlation %>% mean()
sample1.avg.cor.sd <- cor.df$correlation %>% sd()
print(paste0("Sample 1 Average Correlation : ", sample1.avg.cor, " +- ", sample1.avg.cor.sd))
```

```{r}
ggplot(cor.df, aes(x = correlation)) +
  geom_dotplot(binwidth = .01) +
  labs(title = "Sample 1 Cluster Averages Technical Replication", y = element_blank(), x = "Pearson Correlation") +
  coord_flip() +
  geom_text_repel(aes(y = .001, x = correlation, label=ifelse(correlation < .9, cell.type, '')), label.padding = 5, force = 2) 
```

```{r}
#ggsave(paste0(results_dir, "kc_40_technical_cluster_cor_", Sys.Date(), ".png"))
```

```{r}
kc.42 <- subset(seu, subset = biorep == "kc.42")
kc.42$orig.ident <- factor(kc.42$orig.ident, labels = c("Sample 2A", "Sample 2B"))

umap.sample2 <- DimPlot(kc.42, group.by = "orig.ident", pt.size = umap.pt.size) + ggtitle("") + theme(legend.text = element_text(size = legend.text.size))

DimPlot(kc.42, group.by = "orig.ident")
#ggsave(paste0(results_dir, "seu_group_by_kc42_technical_legend_", Sys.Date(), ".png"))
rm(kc.42)
```

```{r}
# Subset to the technical replicate
kc.42.1 <- subset(seu, subset = orig.ident == "kc.42.1") %>% NormalizeData()
# Average expression by cell type
expr.1a <- AverageExpression(kc.42.1, group.by = "cell.type")
# Clean-up the returned object
a2 <- expr.1a$RNA %>% as.data.frame()
a2$`Fetal Cytotrophoblasts` <- NULL

kc.42.2 <- subset(seu, subset = orig.ident == "kc.42.2") %>% NormalizeData()
expr.1b <- AverageExpression(kc.42.2, group.by = "cell.type")
b2 <- expr.1b$RNA %>% as.data.frame()

# Subset a1 and b1 to the cell types that are contained in both
common <- intersect(a2 %>% colnames, b2 %>% colnames)

a2 <- a2 %>% dplyr::select(all_of(common))
b2 <- b2 %>% dplyr::select(all_of(common))

corr.test.res <- map2(a2,
                      b2,
                      ~ cor.test(.x, .y, method = "pearson"))

corr.test.pvals <- map(corr.test.res,
                   ~ .x['p.value'])

pvals <- flatten(corr.test.pvals) %>% flatten

corr.res <- map2(a2,
     b2,
     ~ cor(.x, .y))

cor.df <- corr.res %>% data.frame %>% t
colnames(cor.df) <- "correlation"
cor.df <- as.data.frame(cor.df)

cor.df$cell.type <- rownames(cor.df)
```

```{r}
sample2.avg.cor <- cor.df$correlation %>% mean()
sample2.avg.cor.sd <- cor.df$correlation %>% sd()
print(paste0("Sample 2 Average Correlation : ", sample2.avg.cor, " +- ", sample2.avg.cor.sd))
```

```{r}
ggplot(cor.df, aes(x = correlation)) +
  geom_dotplot(binwidth = .01) +
  labs(title = "Sample 2 Cluster Averages Technical Replication", y = element_blank(), x = "Pearson Correlation") +
  coord_flip() +
  geom_text_repel(aes(y = .001, x = correlation, label=ifelse(correlation < .9, cell.type, '')), label.padding = 5, force = 2) 
```

```{r}
#ggsave(paste0(results_dir, "kc_42_technical_cluster_cor_", Sys.Date(), ".png"))
```


```{r}
panel <- ggarrange(umap.sample1, umap.sample2, umap.biorep, umap.mf, ncol = 2, nrow = 2, labels = "AUTO", font.label = list(size = 30, face = "bold", color = "black"))
ggexport(panel, filename = here("results", "seurat", "umap", paste0(Sys.Date(), "_umaps_var.png")), width = 1980, height = 1080)
```

```{r}
seu$cell.type %>% factor %>% levels
```

```{r}
table <- table(seu$orig.ident, seu$cell.type)
```


```{r}
table <- table(Idents(seu), seu$ident)
samples <- c("1A", "1B", "2A", "2B", "3", "4", "5", "6", "7", "8C", "8P", "9C", "9P")
colnames(table) <- samples
table.df <- as.data.frame(table)


#table.df <- table.df %>% rownames_to_column(var = "cell.type") #%>%
#  arrange(cell.type) #%>%
#  column_to_rownames(var = "cell.type")

total <- sum(table.df$Freq) %>% as.numeric

#write.csv(table, file = here("results", "seurat", paste0(Sys.Date(), "_cell_counts.csv")))
```

### Compare proliferative and non-proliferative cytotrophoblasts - Supplementary Figure

Compare proliferative and cytotrophoblasts.
```{r}
prolif.genes <- FindMarkers(seu, ident.1 = "Fetal Proliferative Cytotrophoblasts", "Fetal Cytotrophoblasts")
```

Full list of genes, formatted for publication
```{r}
p_val_adj_threshold <- 0.05

prolif.markers <- prolif.genes %>%
  rownames_to_column(var = "gene") %>%
  #filter(avg_log2FC > 0) %>%
  filter(p_val_adj < p_val_adj_threshold) %>%
  #group_by(cluster) %>%
  arrange(desc(abs(avg_log2FC))) %>%
  #slice_head(n=10) %>%  # Formatting after this line
  mutate(p_val = if_else(p_val == 0, 2.23e-308, p_val)) %>%
  mutate(p_val_adj = if_else(p_val_adj == 0, 2.23e-308, p_val_adj)) %>%
  dplyr::select(gene, avg_log2FC, pct.1, pct.2, p_val, p_val_adj) %>%
  mutate(p_val = format(p_val, scientific = TRUE, digits = 3)) %>%
  mutate(p_val_adj = format(p_val_adj, scientific = TRUE, digits = 3)) %>%
  mutate(avg_log2FC = format(avg_log2FC, digits = 3)) %>%
  mutate('p-value' = if_else(p_val == 2.23e-308, "< 2.23E-308", as.character(p_val))) %>%
  mutate('adjusted p-value' = if_else(p_val_adj == 2.23e-308, "< 2.23E-308", as.character(p_val_adj))) %>%
  dplyr::select(-c(p_val, p_val_adj))

prolif.markers.colnames <- c("Gene", "Average log2 Fold-change", "Percentage cluster cells expressing gene", "Percentage other cells expressing gene", "P-value", "Bonferroni-adjusted p-value")
colnames(prolif.markers) <- prolif.markers.colnames

# Write to .csv
#write.csv(prolif.markers, paste0(here("results", "seurat", paste0("/", Sys.Date(),  "prolif_ct_markers.csv"))), row.names = FALSE)
```

```{r}
prolif.markers$Gene %>% length
```

```{r}
# Function to select the gene column in each sublist
marker_Volcano <- function(res, names, num.to.plot, genes.of.interest, threshold, output_dir) {
  
  # Indicator variable as to whether significant
  res$sig <- (res$p_val_adj < threshold)
  
  res <- rownames_to_column(res, var = "gene")
  
  # Get the top downregulated hits
  downreg <- res %>%
    filter(p_val_adj < threshold) %>%
    filter(avg_log2FC < 0) %>%
    arrange(avg_log2FC)
  #Get the top upregulated hits.
  upreg <- res %>%
    filter(p_val_adj < threshold) %>%
    filter(avg_log2FC > 0) %>%
    arrange(desc(avg_log2FC))
  
  top.upreg <- upreg$gene[1:num.to.plot]
  print(top.upreg)
  top.downreg <- downreg$gene[1:num.to.plot]
  genes.to.plot <- c(top.upreg, top.downreg, genes.of.interest)
  print(top.downreg)
  res$label <- F
  res$label[(res$gene %in% genes.to.plot)] <- T
  
  p <- ggplot(res) +
    geom_point(aes(
      x = avg_log2FC,
      y = -log10(p_val_adj),
      colour = sig
    )) +
    geom_text_repel(aes(
      x = avg_log2FC,
      y = -log10(p_val_adj),
      label = ifelse(label, gene, "")
    )) +
    geom_hline(yintercept = -log10(threshold),
               linetype = "dotted") +
    ggtitle(names) +
    xlab("log2 fold change") +
    ylab("-log10 adjusted p-value") +
    theme(
      legend.position = "none",
      plot.title = element_text(size = rel(1.5), hjust = 0.5),
      axis.title = element_text(size = rel(1.25))
    )
  print(p)
  ggsave(plot = p, filename = paste0(output_dir, names, "_", Sys.Date(), ".png"), device = "png")
}
```

```{r}
marker_Volcano(prolif.genes, "Proliferative vs. Non-Proliferative Cytotrophoblasts", num.to.plot = 10, genes.of.interest = NULL, threshold = 0.05, output_dir = here("results", "seurat"))
```

```{r}
list.gProfiler <- prolif.genes %>% rownames_to_column(var = "gene") %>% arrange(desc(avg_log2FC)) %>% dplyr::select(gene) 
list.gProfiler <- list.gProfiler$gene
gostres <- gost(query = list.gProfiler,
               organism = "hsapiens", ordered_query = TRUE,
               multi_query = FALSE, significant = TRUE, exclude_iea = TRUE,
               measure_underrepresentation = FALSE, evcodes = FALSE,
               user_threshold = 0.05, correction_method = "g_SCS",
               domain_scope = "annotated", custom_bg = NULL,
               numeric_ns = "", sources = "GO:BP", as_short_link = FALSE)

# Code snippet to modify the term_name label to fit plotting window later
gostres$result <- gostres$result %>%
  mutate(ifelse(term_name == "L13a-mediated translational silencing of Ceruloplasmin expression", "L13a translational silencing of Ceruloplasmin expression", term_name))
```

Get full enrichment results
```{r}
# Drop parent column list for .csv save of all results
gostres.table <- gostres$result
gostres.table$parents <- NULL

gostres.table <- 
  gostres.table %>%
  relocate(term_name, .after = significant) %>%
  dplyr::select(-c(query)) 
#write.csv(gostres.table, here("results", "seurat", paste0("prolif_ct_vs_ct_markers_gost_results_table_", Sys.Date(), ".csv")), row.names = FALSE)
```

```{r}
p <- gostplot(gostres, capped = FALSE, interactive = FALSE) + ggtitle("Proliferative vs. Non-Proliferative Cytotrophoblasts")
p
```

```{r}
# Drop parent column list for .csv save of all results
gostres.table <- gostres$result
gostres.table$parents <- NULL

gostres.table <- 
  gostres.table %>%
  relocate(term_name, .after = significant)
#write.csv(gostres.table, here("results", "seurat", paste0("seurat_all_markers_gost_results_table_", Sys.Date(), ".csv")), row.names = FALSE)

gostres.table$query %>% factor %>% levels

# Drop parent column list for .csv save of filtered GO:BP results
#gostres.small.table <- gostres.small$result
#gostres.small.table$parents <- NULL

#gostres.small.table <- gostres.small.table %>%
  #dplyr::filter(source == "GO:BP") %>%
  #dplyr::relocate(term_name, .after=query) %>%
  #dplyr::select(!significant)


cluster.ontologies <- gostres.table %>% #gostres.small.table %>%
  dplyr::select(term_name, p_value, term_size, query_size, intersection_size, source, term_id) %>%
  arrange(p_value)
colnames(cluster.ontologies) <- c("Term name", "Adjusted P-value", "Term size", "Query size", "Intersection size", "Ontology Source", "Term ID")
```

```{r}
#terms.to.plot <- gostres$result %>% group_by(source) %>% filter(!(source %in% c("HPA", "MIRNA", "CORUM" , "TF", "GO:CC", "WP"))) %>% slice_min(order_by = p_value, n = 3)
terms.to.plot <-
  gostres$result %>%
  filter(source %in% c("GO:BP")) %>%
  #filter(!(source %in% c("HPA", "MIRNA", "CORUM" , "TF", "WP"))) %>%
  arrange(p_value) %>%
  #filter(!(term_id == "REAC:R-HSA-156827")) %>%
  filter(term_size < 500 & term_size > 15)
  #filter(term_size < 1000 & term_size > 125)
  
#terms.to.plot <- gostres$result %>% filter(source == "GO:BP") %>% arrange(p_value)
#terms.to.plot <- terms.to.plot$term_id[1:10]
terms.to.plot <- terms.to.plot$term_id[1:10]
```

```{r}
#pp <- publish_gostplot(p, highlight_terms = terms.to.plot, filename = here("results", "seurat", paste0(Sys.Date(), "_prolif_ct_enriched_gost_plot.png")))
```

### Export proliferative vs. cytotrophoblast results to multiple sheets
```{r}
list <- list(dex = prolif.markers, gost = cluster.ontologies)
names(list) <- c("Differential Expression", "Gene Ontology Enrichment")
#write.xlsx(x = list, file = here("results", "seurat", paste0(Sys.Date(), "_prolif_ct_vs_ct_res.xlsx")))
```

Compare mesenchymal stem cells and fibroblasts
```{r}
prolif.genes <- FindMarkers(seu, ident.1 = "Fetal Mesenchymal Stem Cells", ident.2 = "Fetal Fibroblasts")
```

```{r}
View(prolif.genes %>% rownames_to_column(var = "gene"))
```

```{r}
marker_Volcano(prolif.genes, "Mesenchymal Stem Cells vs. Fibroblasts", num.to.plot = 10, genes.of.interest = NULL, threshold = 0.05, output_dir = here("results", "seurat"))
```

```{r}
list.gProfiler <- prolif.genes %>% rownames_to_column(var = "gene") %>% arrange(desc(avg_log2FC)) %>% dplyr::select(gene) 
list.gProfiler <- list.gProfiler$gene
gostres <- gost(query = list.gProfiler,
               organism = "hsapiens", ordered_query = TRUE,
               multi_query = FALSE, significant = TRUE, exclude_iea = TRUE,
               measure_underrepresentation = FALSE, evcodes = FALSE,
               user_threshold = 0.05, correction_method = "g_SCS",
               domain_scope = "annotated", custom_bg = NULL,
               numeric_ns = "", sources = NULL, as_short_link = FALSE)
```

```{r}
p <- gostplot(gostres, capped = FALSE, interactive = TRUE) #+ ggtitle("Mesenchymal Stem Cells vs. Fibroblasts")
p
```

```{r}
terms.to.plot <- gostres$result %>% arrange(p_value)
terms.to.plot <- terms.to.plot$term_id[1:10]
```

```{r}
#pp <- publish_gostplot(p, highlight_terms = terms.to.plot, filename = here("results", "seurat", paste0(Sys.Date(), "_mesenchymal_stem_vs_fibroblast_enriched_gost_plot.png")))
```
